```{r}
data <- read.csv("data_aal.csv")
#data <- read.csv("data_ba.csv")
#data <- read.csv("data_lmt.csv")
#data <- read.csv("data_sp.csv")
##data <- read.csv("data_united.csv")
data <- read.csv("data_delta.csv")
head(data)
```

balanced corplot is much better, and lm indicators are also much better, but no other predictors are relavent other that slash
```{r}
imbalanced_predictors = c("moon", "slash", "plus", "star", "two", "lunar", "solar", "flare")

# GPT generated code
majority_class <- subset(data, rowSums(data[imbalanced_predictors]) == 0)
minority_class <- subset(data, rowSums(data[imbalanced_predictors]) > 0)

undersampled_majority <- majority_class[sample(nrow(majority_class), nrow(minority_class), replace = FALSE), ]

# Combine undersampled majority class and minority class
balanced_data <- rbind(undersampled_majority, minority_class)
head(balanced_data, 10)
# lets see how the data is distributed now in our balanced dataset and shuffle it 
bal_data <- balanced_data[sample(nrow(balanced_data)), ] # change to data if you want to use

```


```{r}
summary(data)
```



```{r}
plot_histograms_separate <- function(data, columns) {
  for (col in columns) {
    hist(data[[col]], main = col, xlab = col)
  }
}

selected_columns <- c("Volume" ,"change_cat", "change", "moon", "slash", "plus", "star", "two", "lunar", "solar", "flare")

plot_histograms_separate(data, selected_columns)

```

```{r}
n_rows <- nrow(data)

train_proportion <- 0.8

n_train <- round(train_proportion * n_rows)

train_indices <- sample(1:n_rows, n_train, replace = FALSE)

train_data <- data[train_indices, ]

test_data <- data[-train_indices, ]
```


```{r}
model <- lm(change ~ moon + slash + plus + star + two + flare+lunar+solar, data = train_data)

cat("r2",summary(model)$r.squared)
summary(model)
```



```{r}
plot(model$residuals, main = "Residuals Plot", xlab = "Index", ylab = "Residuals")
```

As we see there is no non-linear relationship in the residuals, so running a non-linear model would not be beneficial.

```{r}

plot(fitted(model), residuals(model), main = "fitted v residuals", xlab = "fitted", ylab = "residuals")

```

```{r}
predictions <- predict(model, newdata = test_data)

# Calculate Mean Squared Error (MSE)
mse <- mean((test_data$change - predictions)^2)
mse
```


Trying the model with just the slash column since it is only valid predictor
```{r}
model <- lm(change ~ slash, data = train_data)

cat("r2",summary(model)$r.squared)
# Print the summary of the model
summary(model)

```


```{r}
predictions <- predict(model, newdata = test_data)

mse <- mean((test_data$change - predictions)^2)
mse
```
We have very similar mse to the one with all of the predictors.



# tring to remove outliers, shows that outliers are the slash predictors
```{r}
df_filtered <- data[!(abs(data$change) > 1.25), ]
```

```{r}
n_rows <- nrow(df_filtered)

train_proportion <- 0.8

n_train <- round(train_proportion * n_rows)

train_indices <- sample(1:n_rows, n_train, replace = FALSE)

train_data <- df_filtered[train_indices, ]

test_data <- df_filtered[-train_indices, ]
```


```{r}
model <- lm(change ~ moon + slash + plus + star + two + flare+lunar+solar, data = train_data)

# Print the summary of the model
summary(model)
```

```{r}
predictions <- predict(model, newdata = test_data)

# Calculate Mean Squared Error (MSE)
mse <- mean((test_data$change - predictions)^2)
mse
```

```{r}
plot(model$residuals, main = "Residuals Plot", xlab = "Index", ylab = "Residuals")
```
Removing the outliers from the data made the residuals spread even better, but did not help with model

```{r}
model <- glm(change_cat ~ moon + slash + plus + star + two + flare+lunar+solar, data = train_data, family = binomial)

# Summary of the model
summary(model)
```


```{r}
predictions <- predict(model, newdata = test_data, type = "response")

predicted_labels <- factor(ifelse(predictions > 0.5, 1, 0), levels = c(1, 0))

actual_labels <- factor(test_data$change_cat, levels = c(1, 0))

conf_matrix <- confusionMatrix(predicted_labels, actual_labels)
print(conf_matrix)
```

```{r}
filtered_df <- data[data$change > 1.25, ]
length(filtered_df)
# Plot the filtered dataframe
plot(filtered_df$X, filtered_df$change, type = "o", col = "blue", 
     xlab = "x", ylab = "y", main = paste("Rows with value", 1))
```

```{r}
library(corrplot)
data_excluded <- filtered_df[, !(names(data) %in% "Date")]
data_excluded <- data_excluded[, !(names(data_excluded) %in% "High")]
data_excluded <- data_excluded[, !(names(data_excluded) %in% "Low")]
data_excluded <- data_excluded[, !(names(data_excluded) %in% "X")]
data_excluded <- data_excluded[, !(names(data_excluded) %in% "cat_change")]

correlation_matrix <- cor(data_excluded)
corrplot(correlation_matrix)
```

is slash causing the outliers??
```{r}
data$outlier <- ifelse(data$change > 1, 1, 0)
```

```{r}
n_rows <- nrow(data)

train_proportion <- 0.8

n_train <- round(train_proportion * n_rows)

train_indices <- sample(1:n_rows, n_train, replace = FALSE)

train_data <- data[train_indices, ]

test_data <- data[-train_indices, ]
```

```{r}
model <- lm(outlier ~ moon + slash + plus + star + two + flare+lunar+solar, data = train_data)

# Print the summary of the model
summary(model)
```

```{r}
data$cat_change <- ifelse(data$change_cat == 1, "1", "0")
```

```{r}
n_rows <- nrow(data)

train_proportion <- 0.8

n_train <- round(train_proportion * n_rows)

train_indices <- sample(1:n_rows, n_train, replace = FALSE)

train_data <- data[train_indices, ]

test_data <- data[-train_indices, ]
```


```{r}
model <- glm(change_cat ~ moon + slash + plus + star + two + flare+lunar+solar, data = train_data, family = binomial)

# Print the summary of the model
print(summary(model))
```

```{r}
predictions <- predict(model, newdata = test_data, type = "response")

predicted_labels <- factor(ifelse(predictions > 0.5, 1, 0), levels = c(1, 0))

actual_labels <- factor(test_data$outlier, levels = c(1, 0))

conf_matrix <- confusionMatrix(predicted_labels, actual_labels)
print(conf_matrix)
```




Random forest model
```{r}
n_rows <- nrow(data)

train_proportion <- 0.8

n_train <- round(train_proportion * n_rows)

train_indices <- sample(1:n_rows, n_train, replace = FALSE)

train_data <- data[train_indices, ]

test_data <- data[-train_indices, ]
```

```{r}
train_data
```

```{r}
# Load the randomForest package
library(randomForest)

# Function to shuffle the data in a custom way
custom_shuffle <- function(data) {
  imbalanced_predictors = c("moon", "slash", "plus", "star", "two", "lunar", "solar", "flare")

  # GPT generated code
  majority_class <- subset(data, rowSums(data[imbalanced_predictors]) == 0)
  minority_class <- subset(data, rowSums(data[imbalanced_predictors]) > 0)
  
  undersampled_majority <- majority_class[sample(nrow(majority_class), nrow(minority_class), replace = FALSE), ]
  
  # Combine undersampled majority class and minority class
  balanced_data <- rbind(undersampled_majority, minority_class)
  head(balanced_data, 10)
  # lets see how the data is distributed now in our balanced dataset and shuffle it 
  bal_data <- balanced_data[sample(nrow(balanced_data)), ] # change to data if you want to use
  return(bal_data)
}

# Train a Random Forest model with custom shuffling
custom_rf_model <- randomForest(change_cat ~ moon + slash + plus + star + two + flare+lunar+solar,
                                data = train_data, 
                                sampsize = nrow(train_data), 
                                strata = custom_shuffle(train_data))

# Make predictions on the entire dataset
predictions <- predict(custom_rf_model, newdata = test_data)

# Evaluate the model
accuracy <- mean(predictions == test_data$change_cat)
print(paste("Accuracy:", accuracy))

```

```{r}
nrow(train_data)
```

```{r}
predictions <- predict(custom_rf_model, newdata = test_data, type = "response")

predicted_labels <- factor(ifelse(predictions > 0.5, 1, 0), levels = c(1, 0))

actual_labels <- factor(test_data$outlier, levels = c(1, 0))

conf_matrix <- confusionMatrix(predicted_labels, actual_labels)
print(conf_matrix)
```





MAKE GRAPH PLOTTING CHANGE, AND LINE WHERE SLASH


penumbral lunar eclipse is slash
Only slash shows up, and only for united and aal, the 2 airline companies, even shows up for delta
